The not so widening achievement gap: an interntational comparison of the evolution of the SES achievement gap from 2000-2015

### 2017-09-11 paper/paper.Rnw
Fixed most graph/table positons - ran everything, so far good

% Try plotting this for each country and wave
% Like, every country as a row and 6 columns for each wave.
\begin{figure}
\begin{center}
<<variance_pisa, include = T, out.height = '3in', out.width = '3.5in', fig.align = 'center'>>=

@@

var_years %>%
  setNames(seq(2000, 2015, 3)) %>%
  enframer("name") %>%
  ggplot(aes(name, value)) +
  geom_col() +
  scale_y_continuous(name = "Standard deviation of raw test scores") +
  scale_x_discrete(name = NULL) +
  ggtitle("Standard deviation of test scores over time") +
  theme_few() +
  theme(text = element_text(size = 14))
@
\caption{Standard deviation of test scores across all waves.}
\end{center}
\end{figure}

@@

\subsection{Descriptives}

@@

\subsection{Evolution of the achievement gap}
% TODO: Calculate the difference between 90/10 and then the ratio of this with the annual gain of one year of schooling (\citet{micklewright} did it).

\begin{figure}
\begin{center}
<<graphing_9010gaps, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
# 90/10 gaps acros countries
diff_increase_fun <- function(df) {

@@

pooled_trendline %>%
  filter(type_test == "math") %>%
  ggplot(aes(as.character(wave), difference)) +
  geom_point() +
  geom_linerange(aes(ymin = 0, ymax = difference))+
  geom_line(data = pooled_trendline, stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 2), size = 0.7,
            colour = "red") +
  facet_wrap(~ country, ncol = 5) +
  scale_y_continuous(name = "90/10 gap in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        panel.grid.major.y = element_line(colour = "grey")) +
  ggtitle("Evolution of the 90/10 achievement gap")
@
\caption{Evolution of the achievement gap for selected countries}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<graphing_perc_increase, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
perc_graph <- function(df, test, title, subtitle = NULL) {
  df %>%
  enframe(name = "country") %>%
  unnest(value) %>%
  filter(country %in% countries) %>%
  select(-country1, -years_available) %>%
  mutate(diff_95_lower = perc_diff - 2*sd_year,
         diff_95_upper = perc_diff + 2*sd_year) %>%
  setNames(c("Country", "Type of test", "Average % difference", "Average SD",
             "Lower 50% bound", "Upper 50% bound", "Lower 95% bound", "Upper 95% bound")) %>%
  filter(`Type of test` == test) %>%
  arrange(`Average % difference`) %>%
  mutate(Country = ordered(forcats::as_factor(Country)),
         Country_num = as.numeric(Country)) %>%
  ggplot(aes(Country, `Average % difference`, fill = `Average % difference` > 0)) +
  geom_hline(yintercept = 0, linetype = "longdash") +
  geom_point() +
  geom_ribbon(aes(x = Country_num, ymin = `Lower 95% bound`, ymax = `Upper 95% bound`), alpha = 0.3) +
  geom_ribbon(aes(x = Country_num, ymin = `Lower 50% bound`, ymax = `Upper 50% bound`), alpha = 0.2) +
  geom_linerange(aes(ymin = 0, ymax = `Average % difference`,
                     colour = `Average % difference` > 0), alpha = 0.4) +
  scale_y_continuous(name = "Avg % increase/decrease",breaks = seq(-160, 160, 40)) +
  scale_fill_discrete(guide = FALSE) +
  scale_colour_discrete(guide = FALSE) +
  coord_cartesian(ylim = c(-160, 160)) +
  ggtitle(title, subtitle) +
  ggthemes::theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

@@

# gridExtra::grid.arrange(m1, m2, m3, nrow = 1)
@
\caption{Percentage change in the achievement gap from 2000 to 2015}
\end{center}
\end{figure}

@@

% Add confidence intervals
\begin{figure}
\begin{center}
<<rate_increase/decrease, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
# You left off here. Plot the gap for 1 and 0 for all countries in the graph before.
complete_data_topbottom %>%
  filter(country %in% countries) %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  select(wave, country, matches("*._math$")) %>%
  gather(ses, gap_size, matches("^\\d_mean_math$")) %>%
  separate(ses, c("ses", "delete"), sep = 2) %>%
  mutate(ses = gsub("_", "", ses)) %>%
  filter(complete.cases(.)) %>%
  # gather(lower_bound, lower, matches("\\d_lower")) %>%
  # gather(upper_bound, upper, matches("\\d_upper"))
  ggplot(aes(as.factor(wave), gap_size, group = ses, colour = ses, shape = ses)) +
  geom_point(size = 2) +
  geom_line() +
  geom_line(stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 1), linetype = "longdash",
            colour = "black") +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the achievement gap by top/bottom groups") +
  # geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1, alpha = 0.4) +
  # geom_hline(yintercept = 0, linetype = "longdash") +
  # coord_cartesian(ylim = c(0, 3)) +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the achievement gap for separate SES groups}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<graphing_allgaps, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=

# Comparing all gaps across countries
complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"),
         type_test == "math") %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  ggplot(aes(as.factor(wave), difference, group = type_test)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_linerange(aes(ymin = 0, ymax = difference)) +
  geom_line(stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 2), size = 0.7,
            colour = "red") +
  geom_point(size = 0.5, alpha = 0.4) +
  scale_y_continuous(name = "90/10 gap in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  coord_cartesian(ylim = c(0, 3)) +
  ggtitle("Comparison of the achievement gap for three socio-economic groups") +
  facet_grid(country ~ type) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"))
@
\caption{Evolution of 90/50/10 SES gaps for selected countries}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<rate_change_graph, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=

@@

full_data %>%
  ggplot(aes(low_increase, high_increase), alpha = 0.2) +
  geom_rect(data = rect_data, aes(xmin = xst,
                                  xmax = xen,
                                  ymin = yst,
                                  ymax = yen),
            fill = rect_data$colour,
            alpha = 0.2,
            inherit.aes = FALSE) +
  geom_line(stat="smooth", method = "lm", se = FALSE, alpha = 0.5, colour = "grey", size = 1,
            linetype = "longdash") +
  geom_point(alpha = 0.2) +
  geom_point(data = filter(full_data, continent == "my_cnt"), colour = "red", alpha = 0.7) +
  geom_line(data = filter(full_data, continent == "my_cnt"), colour = "red", alpha = 0.4,
            stat="smooth", method = "lm", se = FALSE, alpha = 0.5, colour = "red", size = 1) +
  geom_text_repel(data = filter(full_data, continent == "my_cnt"),
                  aes(label = country), box.padding = unit(2.7, "lines")) +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  xlim(lims$xlim) +
  ylim(lims$ylim) +
  coord_cartesian(expand = FALSE) +
  annotate(geom = "text", x = 0.15, y = -0.2,
           label = "Low SES is catching up \n faster than High SES",
           fontface = 2, size = 3) +
  annotate(geom = "text", x = -0.075, y = 0.20,
           label = "High SES is increasing  \n faster than Low SES is catching up",
           fontface = 2, size = 3) +
  labs(x = "Average increase of low SES in SD", y = "Average increase of high SES in SD") +
  theme_minimal()
@
\caption{Rate at which top/bottom SES groups are catching up for the 90/10 SES gap}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<graphing_ses_growth, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
# Graphing how the top/bottom are evolving over time instead of absolute difference

complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"), type_test == "math") %>%
  select(wave, country, type_test, type, contains("math")) %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  gather(score, value, -(wave:type)) %>%
  separate(score, c("ses", "score"), sep = 2) %>%
  spread(score, value) %>%
  ggplot(aes(as.factor(wave), mean_math, group = ses, colour = ses, shape = ses)) +
  geom_line() +
  geom_point(size = 2) +
  coord_cartesian(ylim = c(-1, 2)) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_grid(country ~ type) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Evolution of the SES gaps for top/bottom groups}
\end{center}
\end{figure}

@@

\begin{figure}
\begin{center}
<<avgperformance_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  ggplot(aes(scale(Mean), difference)) +
  geom_smooth(method = "lm") +
  geom_point() +
  scale_y_continuous(name = "90/10 achievement gap") +
  scale_x_continuous(name = "Average performance standardized") +
  ggtitle("Relationship between average performance and achievement gap") +
  facet_wrap(~ wave, nrow = 2) +
  theme_few()
@
\caption{Relationship between 90/10 achievement gap and average performance in mathematics}
\end{center}
\end{figure}

@@

% Add confidence intervals
\begin{figure}
\begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2) +
  geom_line() +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the relationship between achievement and inequality") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
\caption{Time trends of achievement gap and average achievement between 2000 and 2015}
\end{center}
\end{figure}

@@

<<complete_tracking_results, include = TRUE>>=
all_track_models %>%
  stan_table() %>%
  mutate(" " = c(
    "Only 1 track",
    "Age selection >= 15",
    "% of curric tracked",
    "Vocational Index",
    "Intercept"
  )) %>%
  xtable(
    caption = "Explaining 90/10 achievement gap - all countries pooled - Multilevel model"
  ) %>%
  print(
  include.rownames = FALSE,
  add.to.row = table_details(all_track_models)
  )
@

@@

<<graph_complete_tracking_graph, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
new_data <-
  ready_data_age %>%
  with(crossing(country = countries,
                num_tracks = unique(num_tracks),
                age_selection = unique(age_selection),
                zvoc = mean(zvoc, na.rm = TRUE),
                length = typical(length > 0),
                year = "2009"))

@@

<<interaction_plot, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
plot_interaction <-
  marginal_effects(stan_models[[length(stan_models)]],
                 effects = "ztrack:zvoc",
                 int_conditions = list(zvoc = quantile))

@@

order_colour <-
  data_interaction %>%
  distinct(group, colour) %>%
  mutate(colour = rev(colour)) %>%
  pull(colour)

data_interaction %>%
  mutate(linetype = group %>% as.factor,
         colour = factor(colour, levels = order_colour)) %>%
  ggplot(aes(x, y)) +
  geom_line(aes(colour = colour, linetype = linetype), size = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = colour), alpha = 0.4) +
  labs(x = "Tracking Index (SD)", y = "90/10 Achievement gap (SD)") +
  scale_colour_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(100, 0, -25) %>% paste0("%")
  ) +
  scale_linetype_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(100, 0, -25) %>% paste0("%")
  ) +
  scale_fill_discrete(guide = FALSE) +
  ggtitle("Interaction between Tracking and Vocational index")

# pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
# pp_check(mod_tracking, type = "scatter_avg", nsamples = 100)
# pp_check(mod_tracking, type = "boxplot", nsamples = 20)
# pp_check(mod_tracking, type = "error_hist", nsamples = 6)
# pp_check(mod_tracking, type = "ecdf_overlay")
#
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "zvoc")
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "ztrack")
@

@@

<<all_gaps_plot, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center'>>=
model_data_merged <-
  map(all_gaps_models, ~ fixef(.x)[, c(1, 3, 4)] %>% as_tibble) %>%
  setNames(gaps) %>%
  enframe() %>%
  unnest(value) %>%
  mutate_if(is_double, round, 2) %>%
  transmute(model = name,
           terms = all_gaps_models[[1]] %>% fixef %>% row.names %>% rep(3),
           estimate = Estimate,
           lower = `2.5%ile`,
           upper = `97.5%ile`)

model_data_merged %>%
  ggplot(aes(terms, estimate, colour = model)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_point() +
  geom_hline(yintercept = 0) +
  scale_colour_discrete(name = NULL) +
  scale_y_continuous(limits = c(-0.5, 2)) +
  scale_x_discrete(labels = c("Intercept",
                              "Tracking Index",
                              "Vocational * Tracking Index",
                              "Vocational Index")) +
  labs(y = "Estimate from linear model", x = "Predictors") +
  ggtitle("Estimation of tracking/vocational models for different gaps") +
  coord_flip()

# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model
# This linear model explains the pooled 90/10 differences
@

>>>>

% Try plotting this for each country and wave
% Like, every country as a row and 6 columns for each wave.
% \begin{figure}
% \begin{center}
<<variance_pisa, include = T, out.height = '3in', out.width = '3.5in', fig.align = 'center', fig.cap = 'Standard deviation of test scores across all waves'>>=

@@

var_years %>%
  setNames(seq(2000, 2015, 3)) %>%
  enframer("name") %>%
  ggplot(aes(name, value)) +
  geom_col() +
  scale_y_continuous(name = "Standard deviation of raw test scores") +
  scale_x_discrete(name = NULL) +
  ggtitle("Standard deviation of test scores over time") +
  theme_few() +
  theme(text = element_text(size = 14))
@
% \caption{Standard deviation of test scores across all waves}
% \end{center}
% \end{figure}

@@

\subsection{Evolution of the achievement gap}

@@

% TODO: Calculate the difference between 90/10 and then the ratio of this with the annual gain of one year of schooling (\citet{micklewright} did it).

% \begin{figure}
% \begin{center}
<<graphing_9010gaps, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Evolution of the achievement gap for selected countries'>>=
# 90/10 gaps acros countries
diff_increase_fun <- function(df) {

@@

pooled_trendline %>%
  filter(type_test == "math") %>%
  ggplot(aes(as.character(wave), difference)) +
  geom_point() +
  geom_linerange(aes(ymin = 0, ymax = difference))+
  geom_line(data = pooled_trendline, stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 2), size = 0.7,
            colour = "red") +
  facet_wrap(~ country, ncol = 5) +
  scale_y_continuous(name = "90/10 gap in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        panel.grid.major.y = element_line(colour = "grey")) +
  ggtitle("Evolution of the 90/10 achievement gap")
@
% \caption{Evolution of the achievement gap for selected countries}
% \end{center}
% \end{figure}

@@

% \begin{figure}
% \begin{center}
<<graphing_perc_increase, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Evolution of the achievement gap for separate SES groups'>>=
perc_graph <- function(df, test, title, subtitle = NULL) {
  df %>%
  enframe(name = "country") %>%
  unnest(value) %>%
  filter(country %in% countries) %>%
  select(-country1, -years_available) %>%
  mutate(diff_95_lower = perc_diff - 2*sd_year,
         diff_95_upper = perc_diff + 2*sd_year) %>%
  setNames(c("Country", "Type of test", "Average % difference", "Average SD",
             "Lower 50% bound", "Upper 50% bound", "Lower 95% bound", "Upper 95% bound")) %>%
  filter(`Type of test` == test) %>%
  arrange(`Average % difference`) %>%
  mutate(Country = ordered(forcats::as_factor(Country)),
         Country_num = as.numeric(Country)) %>%
  ggplot(aes(Country, `Average % difference`, fill = `Average % difference` > 0)) +
  geom_hline(yintercept = 0, linetype = "longdash") +
  geom_point() +
  geom_ribbon(aes(x = Country_num, ymin = `Lower 95% bound`, ymax = `Upper 95% bound`), alpha = 0.3) +
  geom_ribbon(aes(x = Country_num, ymin = `Lower 50% bound`, ymax = `Upper 50% bound`), alpha = 0.2) +
  geom_linerange(aes(ymin = 0, ymax = `Average % difference`,
                     colour = `Average % difference` > 0), alpha = 0.4) +
  scale_y_continuous(name = "Avg % increase/decrease",breaks = seq(-160, 160, 40)) +
  scale_fill_discrete(guide = FALSE) +
  scale_colour_discrete(guide = FALSE) +
  coord_cartesian(ylim = c(-160, 160)) +
  ggtitle(title, subtitle) +
  ggthemes::theme_few() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

@@

# gridExtra::grid.arrange(m1, m2, m3, nrow = 1)
@
% \caption{Percentage change in the achievement gap from 2000 to 2015}
% \end{center}
% \end{figure}

@@

% Add confidence intervals
% \begin{figure}
% \begin{center}
<<rate_increase/decrease, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Evolution of the achievement gap for separate SES groups'>>=
# You left off here. Plot the gap for 1 and 0 for all countries in the graph before.
complete_data_topbottom %>%
  filter(country %in% countries) %>%
  mutate(country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  select(wave, country, matches("*._math$")) %>%
  gather(ses, gap_size, matches("^\\d_mean_math$")) %>%
  separate(ses, c("ses", "delete"), sep = 2) %>%
  mutate(ses = gsub("_", "", ses)) %>%
  filter(complete.cases(.)) %>%
  # gather(lower_bound, lower, matches("\\d_lower")) %>%
  # gather(upper_bound, upper, matches("\\d_upper"))
  ggplot(aes(as.factor(wave), gap_size, group = ses, colour = ses, shape = ses)) +
  geom_point(size = 2) +
  geom_line() +
  geom_line(stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 1), linetype = "longdash",
            colour = "black") +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the achievement gap by top/bottom groups") +
  # geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1, alpha = 0.4) +
  # geom_hline(yintercept = 0, linetype = "longdash") +
  # coord_cartesian(ylim = c(0, 3)) +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
% \caption{Evolution of the achievement gap for separate SES groups}
% \end{center}
% \end{figure}

@@

% \begin{figure}
% \begin{center}
<<graphing_allgaps, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Evolution of 90/50/10 SES gaps for selected countries'>>=

# Comparing all gaps across countries
complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"),
         type_test == "math") %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  ggplot(aes(as.factor(wave), difference, group = type_test)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_linerange(aes(ymin = 0, ymax = difference)) +
  geom_line(stat = "smooth", method = "lm", aes(group = 1),
            formula = y ~ splines::ns(x, 2), size = 0.7,
            colour = "red") +
  geom_point(size = 0.5, alpha = 0.4) +
  scale_y_continuous(name = "90/10 gap in SD", expand = c(0, 0), lim = c(0, 3)) +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  coord_cartesian(ylim = c(0, 3)) +
  ggtitle("Comparison of the achievement gap for three socio-economic groups") +
  facet_grid(country ~ type) +
  ggthemes::theme_few() +
  theme(panel.spacing = unit(1, "lines"))
@
% \caption{Evolution of 90/50/10 SES gaps for selected countries}
% \end{center}
% \end{figure}

@@

% \begin{figure}
% \begin{center}
<<rate_change_graph, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Rate at which top/bottom SES groups are catching up for the 90/10 SES gap'>>=

@@

full_data %>%
  ggplot(aes(low_increase, high_increase), alpha = 0.2) +
  geom_rect(data = rect_data, aes(xmin = xst,
                                  xmax = xen,
                                  ymin = yst,
                                  ymax = yen),
            fill = rect_data$colour,
            alpha = 0.2,
            inherit.aes = FALSE) +
  geom_line(stat="smooth", method = "lm", se = FALSE, alpha = 0.5, colour = "grey", size = 1,
            linetype = "longdash") +
  geom_point(alpha = 0.2) +
  geom_point(data = filter(full_data, continent == "my_cnt"), colour = "red", alpha = 0.7) +
  geom_line(data = filter(full_data, continent == "my_cnt"), colour = "red", alpha = 0.4,
            stat="smooth", method = "lm", se = FALSE, alpha = 0.5, colour = "red", size = 1) +
  geom_text_repel(data = filter(full_data, continent == "my_cnt"),
                  aes(label = country), box.padding = unit(2.7, "lines")) +
  geom_vline(xintercept = 0, alpha = 0.5) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  xlim(lims$xlim) +
  ylim(lims$ylim) +
  coord_cartesian(expand = FALSE) +
  annotate(geom = "text", x = 0.15, y = -0.2,
           label = "Low SES is catching up \n faster than High SES",
           fontface = 2, size = 3) +
  annotate(geom = "text", x = -0.075, y = 0.20,
           label = "High SES is increasing  \n faster than Low SES is catching up",
           fontface = 2, size = 3) +
  labs(x = "Average increase of low SES in SD", y = "Average increase of high SES in SD") +
  theme_minimal()
@
% \caption{Rate at which top/bottom SES groups are catching up for the 90/10 SES gap}
% \end{center}
% \end{figure}

@@

% \begin{figure}
% \begin{center}
<<graphing_ses_growth, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Evolution of the SES gaps for top/bottom groups'>>=
# Graphing how the top/bottom are evolving over time instead of absolute difference

complete_data_topbottom %>%
  bind_rows(complete_data_topmid) %>%
  bind_rows(complete_data_midbottom) %>%
  filter(country %in% c("United States", "Denmark", "France", "Germany"), type_test == "math") %>%
  select(wave, country, type_test, type, contains("math")) %>%
  mutate(type = factor(type,
                       levels = c("90th/10th SES gap", "90th/50th SES gap", "50th/10th SES gap"),
                       ordered = TRUE)) %>%
  gather(score, value, -(wave:type)) %>%
  separate(score, c("ses", "score"), sep = 2) %>%
  spread(score, value) %>%
  ggplot(aes(as.factor(wave), mean_math, group = ses, colour = ses, shape = ses)) +
  geom_line() +
  geom_point(size = 2) +
  coord_cartesian(ylim = c(-1, 2)) +
  scale_y_continuous(name = "Standardized test scores (mean 0)") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  scale_shape_discrete(name = NULL, labels = c("Low SES", "High SES")) +
  ggtitle("Evolution of the top/bottom groups by different gaps") +
  facet_grid(country ~ type) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
% \caption{Evolution of the SES gaps for top/bottom groups}
% \end{center}
% \end{figure}

@@

% \begin{figure}
% \begin{center}
<<avgperformance_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Relationship between 90/10 achievement gap and average performance in mathematics'>>=
data_to_plot %>%
  ggplot(aes(scale(Mean), difference)) +
  geom_smooth(method = "lm") +
  geom_point() +
  scale_y_continuous(name = "90/10 achievement gap") +
  scale_x_continuous(name = "Average performance standardized") +
  ggtitle("Relationship between average performance and achievement gap") +
  facet_wrap(~ wave, nrow = 2) +
  theme_few()
@
% \caption{Relationship between 90/10 achievement gap and average performance in mathematics}
% \end{center}
% \end{figure}

@@

% Add confidence intervals
% \begin{figure}
% \begin{center}
<<country_achievement_and_inequality, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Time trends of achievement gap and average achievement between 2000 and 2015'>>=
data_to_plot %>%
  mutate(Mean = scale(Mean),
         country = factor(country, levels = ordered_cnt, ordered = TRUE)) %>%
  gather(type, value, difference, Mean) %>%
  filter(country %in% countries) %>%
  ggplot(aes(as.character(wave), value, group = type, colour = type, shape = type)) +
  geom_point(size = 2) +
  geom_line() +
  scale_y_continuous(name = "Test scores in SD") +
  scale_x_discrete(name = NULL, breaks = c(2000, 2009, 2015)) +
  scale_colour_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  scale_shape_discrete(name = NULL, labels = c("Achievement gap", "Avg performance")) +
  ggtitle("Evolution of the relationship between achievement and inequality") +
  facet_wrap(~ country, ncol = 5) +
  theme_few() +
  theme(panel.spacing = unit(1, "lines"),
        legend.position = "bottom")
@
% \caption{Time trends of achievement gap and average achievement between 2000 and 2015}
% \end{center}
% \end{figure}

@@

<<complete_tracking_results, include = TRUE>>=
all_track_models %>%
  stan_table() %>%
  mutate(" " = c(
    "Only 1 track",
    "Age selection >= 15",
    "% of curric tracked",
    "Vocational Index",
    "Intercept"
  )) %>%
  xtable(
    caption = "Explaining 90/10 achievement gap - Multilevel model"
  ) %>%
  print(
  include.rownames = FALSE,
  add.to.row = table_details(all_track_models)
  )
@

@@

<<graph_complete_tracking_graph, include = T, fig.width = 9.5, out.height = '7in', out.width = '6.5in', fig.align = 'center', fig.cap = 'Simulation: predicted 90/10 achievement gap for different tracking setups'>>=
new_data <-
  ready_data_age %>%
  with(crossing(country = countries,
                num_tracks = unique(num_tracks),
                age_selection = unique(age_selection),
                zvoc = mean(zvoc, na.rm = TRUE),
                length = typical(length > 0),
                year = "2009"))

@@

% \begin{figure}[!htbp]
% \begin{center}
<<interaction_plot, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'Interaction between tracking and vocational values. Legend ordered relative to the lines'>>=
plot_interaction <-
  marginal_effects(stan_models[[length(stan_models)]],
                 effects = "ztrack:zvoc",
                 int_conditions = list(zvoc = quantile))

@@

order_colour <-
  data_interaction %>%
  distinct(group, colour) %>%
  mutate(colour = colour) %>%
  pull(colour)

data_interaction %>%
  mutate(linetype = factor(group, levels = 5:1),
         colour = factor(colour, levels = order_colour)) %>%
  ggplot(aes(x, y)) +
  geom_line(aes(colour = colour, linetype = linetype), size = 1) +
  geom_ribbon(aes(ymin = ymin, ymax = ymax, fill = colour), alpha = 0.4) +
  labs(x = "Tracking Index (SD)", y = "90/10 Achievement gap (SD)") +
  scale_colour_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(0, 100, 25) %>% paste0("%")
  ) +
  scale_linetype_discrete(
    name = "Vocational index \n quantiles",
    labels = seq(0, 100, 25) %>% paste0("%")
  ) +
  scale_fill_discrete(guide = FALSE) +
  ggtitle("Interaction between Tracking and Vocational index")

# pp_check(mod_tracking, type = "dens_overlay", nsamples = 100)
# pp_check(mod_tracking, type = "scatter_avg", nsamples = 100)
# pp_check(mod_tracking, type = "boxplot", nsamples = 20)
# pp_check(mod_tracking, type = "error_hist", nsamples = 6)
# pp_check(mod_tracking, type = "ecdf_overlay")
#
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "zvoc")
# pp_check(mod_tracking, type = "error_scatter_avg_vs_x", x = "ztrack")
@
% \caption{Interaction between tracking and vocational values. Legend ordered relative to the lines.}
% \end{figure}
% \end{center}

@@

% \begin{figure}[!htbp]
% \begin{center}
<<all_gaps_plot, include = T, out.height = '5in', out.width = '5.5in', fig.align = 'center', fig.cap = 'What explains achievement gaps? Model comparisons'>>=
model_data_merged <-
  map(all_gaps_models, ~ fixef(.x)[, c(1, 3, 4)] %>% as_tibble) %>%
  setNames(gaps) %>%
  enframe() %>%
  unnest(value) %>%
  mutate_if(is_double, round, 2) %>%
  transmute(model = name,
           terms = all_gaps_models[[1]] %>% fixef %>% row.names %>% rep(3),
           estimate = Estimate,
           lower = `2.5%ile`,
           upper = `97.5%ile`)

model_data_merged %>%
  ggplot(aes(terms, estimate, colour = model, shape = model)) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.1) +
  geom_point(size = 2.5) +
  geom_hline(yintercept = 0) +
  scale_y_continuous(limits = c(-0.5, 2)) +
  scale_x_discrete(labels = c("Intercept",
                              "Tracking Index",
                              "Vocational * Tracking Index",
                              "Vocational Index")) +
  scale_colour_discrete(name = NULL) +
  scale_shape_discrete(name = NULL) +
  labs(y = "Estimate from linear model", x = "Predictors") +
  ggtitle("Estimation of tracking/vocational models for different gaps") +
  coord_flip()

# It under predicts the y value rather than over predict
# It has problems predicting the central tendency of the model
# This linear model explains the pooled 90/10 differences
@
% \caption{What explains achievement gaps? Model comparisons}
% \end{figure}
% \end{center}
